---
title: "HSF Exercise"
author: "Brian J. Smith"
date: "18 March 2025"
output: 
  html_document:
    theme: default
    css: "../exercises.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

Let's apply what we've learned about HSFs. Refer to the lecture slides and the walkthrough code (`03_HSF_walkthrough.R`) as a refresher. Also, use <a href="https://conservancy.umn.edu/bitstream/handle/11299/218272/AppA_HSF_examples.html" target="_blank" rel="noopener noreferrer">Appendix A of Fieberg et al. (2021)</a> as a guide for your code.

I encourage you to use your own data for this exercise. I will be available while you are working to answer any questions.

If you don't have your own data, you can use example data built into `amt`. You can access the fisher data and some covariate rasters like this:

```{r amt_fisher, eval = FALSE}
# Load `amt`
library(amt)

# Location data as 'track_xyt'
dat <- amt_fisher

# Habitat data as list
hab <- amt_fisher_covar

```

We are not covering how to model multiple animals today, so for now,
subset your data (or the fisher data) to just one animal. *Don't use Lupe!* Her data were used in Appendix A of Fieberg et al. (2021). Choose one of the other animals, and we can discuss how (or if) they differ from Lupe.

## Instructions

1. Load your data or the `amt` fisher data. Subset to a single individual.

    * Recall that HSFs assume each point is independent and that all habitat is available.
    * If you have high temporal resolution location data, you might want to subset to a coarser temporal resolution to maintain independence.

2. Before you get started, consider your habitat variables. 

    * Decide *a priori* whether each variable represents a resource, risk, or condition. 
    * Make sure you distinguish between continuous and categorical (like landcover) variables.
    * Include your thoughts about each variable as comments in your analysis script.
    
3. Fit a third-order HSF. *I.e.*, sample available points from within the home range of your individual.

    * Make sure the form of each variable matches your *a priori* hypotheses about it from above (*i.e.*, are your continuous conditions modeled as parabolas?).
    * Use the logistic regression approach by sampling a large number of available locations. Make sure to give the used locations a weight of 1 and the available locations large weights.
    
4. Plot the RSS (or log-RSS) for the range of each of your habitat variables.

    * What does each plot tell you about the biology of your system?
    * Be sure to include confidence intervals.


<br>
<hr>
<br>