---
title: "iSSF Exercise"
author: "Brian J. Smith"
date: "18 March 2025"
output: 
  html_document:
    theme: default
    css: "../exercises.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

Let's apply what we've learned about iSSFs. Refer to the lecture slides and the walkthrough code (`02_SSF_walkthrough.R`) as a refresher. Also, use <a href="https://conservancy.umn.edu/bitstream/handle/11299/218272/AppB_SSF_examples.html" target="_blank" rel="noopener noreferrer">Appendix B of Fieberg et al. (2021)</a> as a guide for your code.

I encourage you to use your own data for this exercise. I will be available while you are working to answer any questions.

If you don't have your own data, you can use some sample cougar data from UT. These data were published by Mahoney et al. (2017) [here](https://www.datarepository.movebank.org/handle/10255/move.590). These data were originally used in this manuscript:

> Mahoney PJ, Young JK (2016) Uncovering behavioural states from animal activity and site fidelity patterns. Methods in Ecology and Evolution 8(2): 174–183. [doi:10.1111/2041-210X.12658](https://doi.org/10.1111/2041-210X.12658)

I've already downloaded and processed the data, which you can find in `data/coyote_cougar.csv`. I've also downloaded and processed some environmental raster data for you to use, which you can find as a multi-band GeoTiff in `data/coyote_cougar_habitat.tif`.

Habitat layers are:  

- Elevation (m)
- Tree cover (%)
- Herbaceous biomass (kg/ha)
- Distance to nearest road (m)

We are not covering how to model multiple animals today, so for now,
subset your data (or the coyote and cougar data) to just one animal.

```{r data, eval = FALSE}
# Load `tidyverse` and `terra`
library(tidyverse)
library(terra)

# Location data
dat <- read_csv("data/coyote_cougar.csv") %>% 
  # Subset to just cougar F53
  filter(id == "F53")

# Set the timezone
tz(dat$t_) <- "US/Mountain"

# Habitat data as stack
hab <- rast("data/coyote_cougar_habitat.tif")
names(hab) <- c("elevation", "trees", "biomass", "dist_to_road")

```

## Instructions

1. Load your data or the example data. Subset to a single individual.

    * Recall that iSSFs assume a constant step duration. Make sure your data are (within a reasonable tolerance) regularly sampled before you continue.
    * You will want a fairly large sample size as your model grows in parameters by including interaction terms.
    
2. Before you get started, consider your habitat variables and your movement variables. 

    * Decide *a priori* whether each variable represents a resource, risk, or condition.
    * Should any of your covariates affect the movement process? 
    * Include your thoughts about each variable as comments in your analysis script.
    
3. Fit an iSSF, with at least one interaction.

    * Consider an interaction between habitat variables to model tradeoffs.
    * Consider an interaction with time of day to model temporal dynamics.
    * Consider an interaction between the movement parameters and a habitat variable to model how habitat affects the movement process.
    
4. Make at least one plot showing the RSS (or log-RSS) for one of your habitat variables.

    * What does this plot tell you about the biology of your system?
    
5. Make at least one plot showing the step-length or turn-angle distribution for your animal.

    * If you have interactions with step-length or turn-angle, try to show this.
    

<br>
<hr>
<br>


